---
title: "homework_3"
format: html
---

[View repository on GitHub](https://github.com/mconrad23/ENVS-193DS_homework-03.git)

#Set up

```{r} 
#load in packages
library(tidyverse) 
library(here)
library(gt) 
library(janitor)
```
#Problem 1. Personal data 

##a. Data summarizing 
[To summarize my data in order to compare the response variable between categories, I should calculate the mean length of study sessions for each type of music to compare how different music genres might affect how long I am studying. This comparison would be informative because the type of music may influence focus and productivity, potentially leading to longer or shorter study sessions depending on the genre.]

##b. Visualization 
```{r}
#load in libraries
library(ggplot2)
library(readxl)
library(dplyr)

# read in excel file 
data <- read_csv("personal_data_project_DATA_Sheet1.csv")

# summarize data - mean study length per music genre
summary_data <- data %>%
  group_by(`Type of Music`) %>%
  summarise(Mean_Length = mean(`Length of Study Session (hours)`, na.rm = TRUE))

# actual plot
ggplot(data, aes(x = `Type of Music`, y = `Length of Study Session (hours)`, color = `Type of Music`)) +
  geom_jitter(width = 0.2, alpha = 0.6, size = 2) +  # show individual study sessions
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, color = "black") +  # mean points
  scale_color_brewer(palette = "Set2") +  # colorful palette
  labs(
    title = "Study Session Lengths by Music Genre",
    x = "Type of Music",
    y = "Length of Study Session (hours)",
    color = "Music Genre"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

##c. Caption 

[Figure 1.The length of each study session in hours by the type of music listened to during each session. Each point represents the length (in hours) of an individual study session under a specific type of music. Colored jittered points show the distribution of individual sessions, while black diamonds represent the mean study length for each genre. This visualization highlights how different music types may be associated with longer or shorter average study duration, which may have effects on productivity and how much work is accompliosged duirng each study session.]

##d. Table presentation 

```{r}
# load extra library
library(gt)

# summarize data (mean study session length by type of music)
summary_table <- data %>%
  group_by(`Type of Music`) %>%
  summarise(`Mean Study Length (hours)` = round(mean(`Length of Study Session (hours)`, na.rm = TRUE), 1))

# create table 
summary_table %>%
  gt() %>%
  tab_header(
    title = "Average Study Session Length by Music Genre"
  ) %>%
  cols_label(
    `Type of Music` = "Type of Music",
    `Mean Study Length (hours)` = "Mean Length (hours)"
  ) %>%
  fmt_number(
    columns = `Mean Study Length (hours)`,
    decimals = 1
  )

```

#Problem 2. Affective visualization 

##a. Describe in words what an affective visualization could look like for your personal data

##b. Create a sketch (on paper) of your idea.

##c. Make a draft of your visualization

##d. Write an artist statement

#Problem 3. Statistical critique

##a. Revisit and summarize

##b. Visual clarity

##c. Aesthetic clarity 

##d. Recommendations


