---
title: "homework_3"
format: html
---

[View repository on GitHub](https://github.com/mconrad23/ENVS-193DS_homework-03.git)

#Set up

```{r} 
#load in packages
library(tidyverse) 
library(here)
library(gt) 
```
#Problem 1. Personal data 

##a. Data summarizing 
[To summarize my data in order to compare the response variable between categories, I calculate the mean length of study sessions for each type of music to compare how different music genres might affect how long I am studying. This comparison would be informative because the type of music may influence focus and productivity, potentially leading to longer or shorter study sessions depending on the genre.]

##b. Visualization 
```{r}
#load in libraries
library(ggplot2)
library(readxl)
library(dplyr)

# Read the Excel file (adjust path if needed)
data <- read_excel("personal_data_project_DATA.xls", sheet = "Sheet1", skip = 1)

# Clean column names
colnames(data) <- trimws(colnames(data))

# Convert study session length to numeric
data$`Length of Study Session (hours)` <- as.numeric(data$`Length of Study Session (hours)`)

# Create the plot
ggplot(data, aes(x = `Type of Music`, y = `Length of Study Session (hours)`)) +
  geom_boxplot(fill = "#A8DADC", color = "#457B9D", outlier.shape = NA) +  # Custom box colors
  geom_jitter(width = 0.2, size = 2, alpha = 0.6, color = "#1D3557") +     # Add individual points
  labs(
    title = "Comparison of Study Session Lengths by Music Type",
    x = "Type of Music Listened to While Studying",
    y = "Length of Study Session (in Hours)"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


##c. Caption 

##d. Table presentation 

#Problem 2. Affective visualization 

##a. Describe in words what an affective visualization could look like for your personal data

##b. Create a sketch (on paper) of your idea.

##c. Make a draft of your visualization

##d. Write an artist statement

#Problem 3. Statistical critique

##a. Revisit and summarize

##b. Visual clarity

##c. Aesthetic clarity 

##d. Recommendations


